---
title: "STAT1200 Seminar 7: Sampling and Estimation"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Introduction

This seminar focuses on topics covered in **Sampling and Estimation**. You will be given time to work through practical examples related to the materials covered in these lectures.

Please feel free to refer to the example code in the lectures to help you solve the problems.

As always, please do not hesitate to ask any questions!

\newpage

## Task 1

Use the following code to import the *diamonds* dataset into R from the *ggplot2* package.

You can take a bit of time to re-familiarize yourself with the data.

#### A

```{r, warning=FALSE}

library(ggplot2)
data(diamonds)
```

```{r}

?diamonds
```

#### B

```{r}

library(BSDA)
library(DescTools)

```

1.  **Assume a known world-wide (population) standard deviation of 4000** ($\sigma=4000$) estimate a 95% confidence interval for the mean of the `price` of the diamonds.

```{r}

```

2.  Estimate a 95% confidence interval for the proportion of diamonds with the `cut` of **Ideal**.

```{r}

```

3.  Assuming that we **do not know** the world-wide standard deviation, estimate a 99% confidence interval for the mean of the `carat` of the diamonds.

```{r}

```

4.  Estimate a 90% confidence interval for the variance of the `carat` of the diamonds.

```{r}

```

5.  Assuming that we **do not know** the world-wide standard deviations, estimate a 90% confidence interval for the differences in the means of the `price` of the diamonds between diamonds with a `cut` of **Ideal** and **Good**. **Comment on your findings.**

```{r}


```

6.  Choose **one** of the first four confidence intervals and briefly comment on what the results imply.

\newpage

## Task 2

#### A

Import the *Wages.csv* file into your R environment and take some time to understand the variables (see Data Description below).

```{r}

Wages = read.csv("Wages.csv")


```

#### B

1.  Estimate a confidence interval for the the differences in proportions of respondents with the `Job` of **Teacher** and **Biologist**. Select $\alpha=0.10$. **Comment on your findings.**

```{r}

```

2.  Assuming that we do **do not know** the nation-wide standard deviations, estimate a confidence interval for the differences in the means between the `Salary` and `Salary.5` salaries scores. Select $\alpha=0.05$. **Comment on your findings.**

```{r}




```

3.  Assuming that we do **do not know** the nation-wide standard deviations, estimate a confidence interval for the differences in the means of the `Salary` between the respondents who live in **rural** and **Metro** `Location`. Select $\alpha=0.01$. **Comment on your findings.**

```{r}


```

4.  Estimate a confidence interval for the variance of the `Age` of the respondents. Select $\alpha=0.05$. **Comment on your findings.**

```{r}

```

5.  Take some time to use confidence intervals to examine different aspects of this sample. Please indicate the significance level you chose.

## Task 3

1.  How many observations $n$, drawn from a random sample will we need to obtain a half confidence interval of 0.66 from a sample with $\sigma = 2.35$?

2.  How many observations $n$, drawn from a random sample will we need to obtain a half confidence interval of 0.5 from a sample with $\sigma = 4.5$?

\newpage

### Data Descriptions

\small

\subsection*{Wages}

This dataset (\textit{Wages.csv}) contains the annual wages and additional information of a sample of 314 respondents from three different Canadian Provinces.

\subsection*{Variables}

\begin{itemize}
\item \textbf{ID:} Unique identifier for each observation.

\item \textbf{Salary:} Respondent's annual salary (thousands of dollars).

\item \textbf{Salary.5:} Respondent's annual salary 5 years prior to the survey.

\item \textbf{Job:} Employment field of the respondent.

\item \textbf{Location:} If the respondent works in a major centre (metro) or a rural area (rural).

\item \textbf{Province:} Province of employment.

\item \textbf{FatherEducation:} The years of education completed by the respondent's father.

\item \textbf{MotherEducation:} The years of education completed by the respondent's mother.

\item \textbf{GPA:} Grade Point Average of the respondent's last completed education.

\item \textbf{Age:} Respondent's age.

\item \textbf{Experience:} Years of work experience.

\item \textbf{Education:} The years of education completed by the respondent.

\end{itemize}
