---
title: "STAT1200 Seminar 7: Sampling and Estimation"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Introduction

This seminar focuses on topics covered in **Sampling and Estimation**. You will be given time to work through practical examples related to the materials covered in these lectures.

Please feel free to refer to the example code in the lectures to help you solve the problems.

As always, please do not hesitate to ask any questions!

\newpage

## Task 1

Use the following code to import the *diamonds* dataset into R from the *ggplot2* package.

You can take a bit of time to re-familiarize yourself with the data.

#### A

```{r, warning=FALSE}

library(ggplot2)
data(diamonds)
```

```{r}

# ?diamonds
```

#### B

Packages:

```{r,message=FALSE}
if(!require(BSDA)) install.packages("BSDA")
library(BSDA)
if(!require(DescTools)) install.packages("DescTools")
library(DescTools)
```

1.  **Assume a known world-wide (population) standard deviation of 4000** ($\sigma=4000$) estimate a 95% confidence interval for the mean of the `price` of the diamonds.

```{r}

pricebar = mean(diamonds$price) # sample mean

# 95% C.I.
# Assume pop. sd is 4000 
Lower.95 <- pricebar - qnorm(p=(0.05/2),lower.tail = FALSE)*
  (4000/sqrt(nrow(diamonds)))
Upper.95 <- pricebar + qnorm(p=(0.05/2),lower.tail = FALSE)*
  (4000/sqrt(nrow(diamonds)))

# OR #

#Using the BSDA package:
#?z.test()
Price.CI <- z.test(diamonds$price,sigma.x = 4000,conf.level = 0.95)
Price.CI$conf.int

```

2.  Estimate a 95% confidence interval for the proportion of diamonds with the `cut` of **Ideal**.

```{r}

#Calculate the sample proportion
Ideal.prop <- sum(as.numeric(diamonds$cut == "Ideal"))/nrow(diamonds)

# 90% C.I.

Ideal.prop.lower.90 <- Ideal.prop - qnorm(p=0.05/2,lower.tail=FALSE)*
  sqrt((Ideal.prop*(1-Ideal.prop))/nrow(diamonds))

Ideal.prop.upper.90 <- Ideal.prop + qnorm(p=0.05/2,lower.tail=FALSE)*
  sqrt((Ideal.prop*(1-Ideal.prop))/nrow(diamonds))


#Or using prop.test()
# ?prop.test()

ideal.prop.test = prop.test(sum(as.numeric(diamonds$cut == "Ideal")),
                            nrow(diamonds), conf.level=0.95) 
ideal.prop.test$conf.int

# or # (same results)

xideal = sum(as.numeric(diamonds$cut == "Ideal")) # number of "Ideal" diamonds
ntotal = nrow(diamonds) # Number of total observations

# Estimate the C.I.
ideal.prop.test2 = prop.test(x=xideal,n=ntotal,conf.level = 0.95) 
ideal.prop.test2$conf.int # Print the interval


```

3.  Assuming that we **do not know** the world-wide standard deviation, estimate a 99% confidence interval for the mean of the `carat` of the diamonds.

```{r}

carat.mean <- mean(diamonds$carat)

n <- nrow(diamonds)
s <- sd(diamonds$carat)
alpha <- 0.01

Carat.Lower.95 <- carat.mean - qt(p=alpha/2, df = n-1, lower.tail = FALSE)*
  (s/sqrt(n))
Carat.Upper.95 <- carat.mean + qt(p=alpha/2, df = n-1, lower.tail = FALSE)*
  (s/sqrt(n))

# OR #

Carat.CI <- t.test(diamonds$carat,conf.level = 0.99)
Carat.CI$conf.int

```

4.  Estimate a 90% confidence interval for the variance of the `carat` of the diamonds.

```{r}

var(diamonds$carat)

VarCI(diamonds$carat,method = "classic",conf.level = 0.90)

```

5.  Assuming that we **do not know** the world-wide standard deviations, estimate a 90% confidence interval for the differences in the means of the `price` of the diamonds between diamonds with a `cut` of **Ideal** and **Good**. **Comment on your findings.**

```{r}

library(DescTools)


MeanDiffCI(diamonds$price[diamonds$cut == "Ideal"],
           diamonds$price[diamonds$cut == "Good"],
           conf.level = 0.90,
  paired = FALSE)


# OR #

# Vectors of each
Ideal = subset(diamonds,diamonds$cut == "Ideal") 
Good = subset(diamonds,diamonds$cut == "Good")

MeanDiffCI(Ideal$price,Good$price,conf.level = .90)

```

6.  Choose **one** of the first four confidence intervals and briefly comment on what the results imply.
    1.  Price: 95% C.I. suggests that with 95% confidence the average price of a diamond is between \$3899.04 and \$3966.56. *This is quite narrow due to the large sample size and the assumption of a known population standard deviation.*
    2.  Ideal Cut proportion: 0.395 to 0.404 suggests that with 95% confidence, the proportion of Ideal cut diamonds is in this range. *Again, narrow due to a large sample size.*
    3.  Carat mean: With 99% confidence the average carets of a diamond is between 0.793 to 0.803. *Did not assume a known population standard deviation.* 
    4.  Carat variance: The variance of the carat of diamonds is between 0.222 and 0.227 with 90% confidence. 
    5.  Difference between average prices Ideal & Good Cuts: **This confidence interval for the difference suggests that with 90% confidence the average price for good diamonds is higher than the average price of ideal cut diamonds.** *Does not account for other attributes.*

\newpage

## Task 2

#### A

Import the *Wages.csv* file into your R environment and take some time to understand the variables (see Data Description below).

```{r}

Wages = read.csv("Wages.csv")


```

#### B

1.  Estimate a confidence interval for the the differences in proportions of respondents with the `Job` of **Teacher** and **Biologist**. Select $\alpha=0.10$. **Comment on your findings.**

```{r}


x.teacher = sum(as.numeric(Wages$Job == "Teacher"))
x.Biologist = sum(as.numeric(Wages$Job == "Biologist"))

n = nrow(Wages)


Job.Prop.diff<- prop.test(x=c(x.teacher,x.Biologist),n=c(n,n),
                          conf.level = 0.90,correct=FALSE)


Job.Prop.diff$conf.int

```

**Comments:** With 90% confidence, there does not appear to be a difference in the proportion of people with jobs as a Teacher or Biologist. **The value 0 is in the 90% confidence interval.**

2.  Assuming that we do **do not know** the nation-wide standard deviations, estimate a confidence interval for the differences in the means between the `Salary` and `Salary.5` salaries scores. Select $\alpha=0.05$. **Comment on your findings.**

```{r}


MeanDiffCI(Wages$Salary,Wages$Salary.5,conf.level = 0.95, paired = TRUE,
           method = "classic")

```

**Comments:** With 95% confidence, the difference is larger than 0 implying that the current salaries are higher than past salaries. **The value 0 is not in the 95% confidence interval.**

3.  Assuming that we do **do not know** the nation-wide standard deviations, estimate a confidence interval for the differences in the means of the `Salary` between the respondents who live in **rural** and **Metro** `Location`. Select $\alpha=0.01$. **Comment on your findings.**

```{r}


MeanDiffCI(Wages$Salary[Wages$Location == "Rural"],
           Wages$Salary[Wages$Location == "Metro"],conf.level = 0.99, 
           paired = FALSE, method = "classic")

```

**Comments:** With 99% confidence, it does not appear that there is a difference in the salaries of people living in Rural and Metro areas. **The value 0 is in the 99% confidence interval for the difference.**

4.  Estimate a confidence interval for the variance of the `Age` of the respondents. Select $\alpha=0.05$. **Comment on your findings.**

```{r}

VarCI(Wages$Age,method = "classic", conf.level = 0.95) # Use this one!!

```

**Comments:** The results indicate that there is moderate variability, but the confidence interval is fairly wide.

5.  Take some time to use confidence intervals to examine different aspects of this sample. Please indicate the significance level you chose.

**I will leave this to you**

## Task 3

1.  How many observations $n$, drawn from a random sample will we need to obtain a half confidence interval of 0.66 from a sample with $\sigma = 2.35$?

```{r,warning=FALSE,message=FALSE}

library(EnvStats)

E <- 0.66
sigma <- 2.35
alpha <- 0.05 #=> C.I. is 95%

n_req <- ((qnorm(p=alpha/2,lower.tail=FALSE)*sigma)/E)^2

n_req

ciNormN(E,sigma,conf.level = 0.95)

```

2.  How many observations $n$, drawn from a random sample will we need to obtain a half confidence interval of 0.5 from a sample with $\sigma = 4.5$?

```{r}

E <- 0.5
sigma <- 4.5
alpha <- 0.05 #=> C.I. is 95%

n_req <- ((qnorm(p=alpha/2,lower.tail=FALSE)*sigma)/E)^2

n_req

ciNormN(E,sigma,conf.level = 0.95)

```

\newpage

### Data Descriptions

\small

\subsection\*{Wages}

This dataset (\textit{Wages.csv}) contains the annual wages and additional information of a sample of 314 respondents from three different Canadian Provinces.

\subsection\*{Variables}

\begin{itemize}
\item \textbf{ID:} Unique identifier for each observation.

\item \textbf{Salary:} Respondent's annual salary (thousands of dollars).

\item \textbf{Salary.5:} Respondent's annual salary 5 years prior to the survey.

\item \textbf{Job:} Employment field of the respondent.

\item \textbf{Location:} If the respondent works in a major centre (metro) or a rural area (rural).

\item \textbf{Province:} Province of employment.

\item \textbf{FatherEducation:} The years of education completed by the respondent's father.

\item \textbf{MotherEducation:} The years of education completed by the respondent's mother.

\item \textbf{GPA:} Grade Point Average of the respondent's last completed education.

\item \textbf{Age:} Respondent's age.

\item \textbf{Experience:} Years of work experience.

\item \textbf{Education:} The years of education completed by the respondent.

\end{itemize}
