---
title: "STAT1200 Seminar 8: Hypothesis Testing"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Introduction

This seminar focuses on topics covered in **Hypothesis Testing I/II/III**. You will be given time to work through practical examples related to the materials covered in these lectures.

Please feel free to refer to the example code in the lectures to help you solve the problems.

As always, please do not hesitate to ask any questions!

\newpage

## Task 1

Use the following code to import the *diamonds* dataset into R from the *ggplot2* package.

You can take a bit of time to re-familiarize yourself with the data.

#### A

```{r, warning=FALSE}

library(ggplot2)
data(diamonds)


?diamonds
```

#### B

1.  It is widely believed that the **average `price`** of diamonds is \$7000. With 95% confidence, does this assumption hold based on your sample? **Be sure to identify the null and alternative hypotheses and make conclusions from your results** (see above example).

```{r}

```

2.  It is widely believed that the **average `depth`** percentage of diamonds is less than or equal to 60. With 95% confidence, does this assumption hold based on your sample? **Be sure to identify the null and alternative hypotheses and make conclusions from your results** (see above example).

```{r}

```

3.  It is assumed that the variance in the `price` of Ideal `cut` diamonds is greater than or equal to the `price` of Premium `cut` diamonds. With 90% confidence, does this assumption hold based on your sample? **Be sure to identify the null and alternative hypotheses and make conclusions from your results.**

```{r}

```

4.  It is widely assumed that the proportion of diamonds with a `color` of D (best) is equal to 0.25. With 99% confidence, does this assumption hold based on your sample? **Be sure to identify the null and alternative hypotheses and make conclusions from your results.**

```{r}


```

## Task 2

**Continuing to work with the diamonds dataset.**

1.  Conduct an appropriate test with 95% significance, to determine if the of the `depth` of the diamonds are the same across all of the `cuts`. Be sure to check all of the necessary assumptions. **Identify the null and alternative hypotheses for each step and be sure to comment on your findings.**

```{r}

```

2.  It is widely believed that the `clarity` of diamonds is independent of their `cut`. Conduct an appropriate hypothesis test to test the validity of this assumption ($\alpha=0.1$). **Identify the null and alternative hypotheses and be sure to comment on your findings.**

```{r}

```

## Task 3

**Continuing to work with the diamonds dataset.**

**Conduct appropriate hypothesis tests to evaluate the following *alternative hypotheses* based on your sample.**

1.  The mean of the `table` for diamonds with a `cut` E is less than the mean of the `table` for diamonds with a `cut` G. Select $\alpha=0.05$. **Be sure to identify the null and alternative hypotheses and make conclusions from your results.**

```{r}
```

2.  The variance of the `carat` of the diamonds is greater than 0.13. Select $\alpha=0.01$. **Be sure to identify the null and alternative hypotheses and make conclusions from your results.**

```{r}
```

3.  The proportion of Premium `cut` diamonds with a `color` of G is greater than the proportion of Ideal `cut` diamonds with a `color` of F. Select $\alpha=0.05$. **Be sure to identify the null and alternative hypotheses and make conclusions from your results.**
