---
title: "STAT1200 Seminar 3: Displaying & Summarizing data"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Introduction

This seminar focuses on topics covered in **Displaying & Summarizing data I** and **Displaying & Summarizing data II**. You will be given time to work through practical examples related to the materials covered in these lectures.

Please feel free to refer to the example code in the lectures to help you solve the problems.

As always, please do not hesitate to ask any questions!

## Task 1

Create a new R markdown or Quarto file to answer the following questions. Be sure to include you name and student number. Set the default *knit* to PDF or Word.

Use this file to complete the rest of the tasks in this seminar.

#### Problem Context:

We have a sample of respondents on a survey regarding salaries and other characteristics. We are interested in characterizing the sample ahead of any further analysis.

\newpage

## Task 2

1.  Import the *Wages.csv* file into your R environment and take some time to understand the variables (see Data Description below).

```{r}

    Wages = read.csv("Wages.csv")


```

2.  Focus your attention on the `Salary` variable and create the following tables and visualizations **and comment on your findings**:

#### 1. Stem-and-Leaf Display:

```{r}

stem(Wages$Salary)

```

\newpage

**Comments:**

-   There appears to be a spike in \$60 000 - \$69 000 salary range.

-   Most of the observations are between \$31 000 and \$159 000.

-   The distribution appears to be slightly positively skewed with a few observations over \$170 000.

-   There are some gaps in the upper tail (higher salaries).

-   Range: \$18 000 - \$206 000

#### 2. Histogram:

```{r}

hist(Wages$Salary,main="Histogram of Salaries",xlab = "Salary")


```

**Comments:**

-   The distribution is **unimodal** with a spike between \$50 000 and \$100 000.

-   Slightly **positively skewed** with a longer right tail (a few very high salaries).

-   Range displayed: \$0 - \$210 000

#### 3. Boxplot:

```{r}

boxplot(Wages$Salary,main = "Boxplot of Salary", ylab="Salary")

```

**Comments:**

-   Identified three possible **outliers** (very high salaries).

-   Slight positive skew, but it is less pronounced in this visualization.

-   In the context of salaries, there appears to be a lot of variability around the medial value.

Hint: *What do each of these methods of displaying a sample show us?\
**Be sure to include appropriate labels.***

3.  Continuing to focus your attention on the `Salary` variable, examine the following numeric characterizations of your sample **and comment on your findings**:

    1.  Sample: mean, median, quartiles, and 10\\% trimmed mean.
        -   Do you notice anything interesting?
    2.  Sample: range, variance, and standard deviation.

\newpage

#### 1.

```{r}

# mean
mean(Wages$Salary)

# median
median(Wages$Salary)

# Quartiles (also gives you the range)
quantile(Wages$Salary)

# 10% trimmed mean
mean(Wages$Salary,trim = .10)


```

**Comments:**

-   The *average* salary in this sample is \$88 535.38

-   The median salary in this sample is \$85 240

-   Mean and median: The mean is slightly higher than median indicating a slight positive skew in the sample.

-   Quantiles: 50% of the people in the sample make between \$65 295 and \$107 155.

-   25% make above \$107 155 and 25% make below \$65 295.

-   The **trimmed mean** is smaller than the mean indicating that there were a few *extreme* values in the high end of the sample (closer to symmetric by removing the highest and lowest observations).

\newpage

#### 2.

```{r}

# Range (not the only way to do this)
range(Wages$Salary)

# Variance
var(Wages$Salary)

# Standard Deviation
sd(Wages$Salary)


```

**Comments:**

-   The range of salaries in this sample appears to be quite large \[\$17 680, \$205 540\].

-   The variance and standard deviations support the conclusion of high variability within our sample of salaries.

\newpage

## Task 3

1.  Continuing with the *Wages.csv* sample but focusing our attention on the `Job` variable, use the following tools to characterize your sample **and comment on your findings**:

    1.  Bar graph (include appropriate labels).

    2.  Sample proportions for each job.

#### Bar graph/Barplot

```{r}

# Remeber we need the table() function to get frequencies

barplot(table(Wages$Job),main = "Barplot of Jobs",
        ylab = "Frequency")



```

**Comments:**

-   There does not appear to be too much disparity in the frequencies across the three groups.

-   Data Scientist is the most represented job in this sample (over 100), followed by Biologist and Teacher.

#### Sample Proportions

```{r}

counts <- table(Wages$Job)
proportions <- prop.table(counts)
proportions

```

**Comments:**

-   The conclusions from the bar plot are supported by the sample proportions.

-   No category has over 37% representation in the sample.

-   Data Scientist has the highest representation (36.94%), followed by Biologist (34.39%) and Teacher (28.66%).

\newpage

2.  Finally, create a comparative box plot for the `salary` variable, split by the `job` variable, include appropriate labels **and comment on your findings.**

#### Comparative Box plot

```{r}

boxplot(Wages$Salary~Wages$Job,
        main = "Comparative Boxplot of Salaries across Jobs",
        ylab = "Salary",xlab = "Job")


```

**Comments:**

-   Biologist has the **largest** variability in this sample.

-   Teacher has the **lowest** variability in this sample.

-   There are two potential outliers identified in the sample.

    -   One in Biologist and one in Data Scientist

-   Biologist appears to be slightly positively skewed.

-   Teacher appears to be slightly negatively skewed.

-   Data Scientist appears to be approximately symmetric.

-   The sample median value of Data Scientist is higher than the other two jobs.

    -   Teacher has the lowest median value.

## Task 4

Render your solutions as a pdf file.

\newpage

\section*{Data Descriptions}

\small

\subsection*{Wages}

This dataset (\textit{Wages.csv}) contains the annual wages and additional information of a sample of 314 respondents from three different Canadian Provinces.

\subsection*{Variables}

\begin{itemize}
\item \textbf{ID:} Unique identifier for each observation.

\item \textbf{Salary:} Respondent's annual salary (thousands of dollars).

\item \textbf{Salary.5:} Respondent's annual salary 5 years prior to the survey.

\item \textbf{Job:} Employment field of the respondent.

\item \textbf{Location:} If the respondent works in a major centre (metro) or a rural area (rural).

\item \textbf{Province:} Province of employment.

\item \textbf{FatherEducation:} The years of education completed by the respondent's father.

\item \textbf{MatherEducation:} The years of education completed by the respondent's mother.

\item \textbf{GPA:} Grade Point Average of the respondent's last completed education.

\item \textbf{Age:} Respondent's age.

\item \textbf{Experience:} Years of work experience.

\item \textbf{Education:} The years of education completed by the respondent.

\end{itemize}
