---
title: "STAT1200 Seminar 5: Random Variables & Discrete & Continuous Probability Distributions"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Introduction

This seminar focuses on topics covered in **Random Variables & Discrete Probability Distributions** and **Continuous Probability Distributions**. You will be given time to work through practical examples related to the materials covered in these lectures.

Please feel free to refer to the example code in the lectures to help you solve the problems.

As always, please do not hesitate to ask any questions!

\newpage

## Task 1

What is the probability distribution (PMF) of the **sum** of one six-sided die and one four-sided die \[1-4\]? This link will explain [four-sided dice](https://en.wikipedia.org/wiki/Four-sided_die)

```{r, warning=FALSE}
library(knitr)



PMF <- data.frame(Results = c(2,3,4,5,6,7,8,9,10),
                  Probability = c(1/24,2/24,3/24,4/24,4/24,4/24,
                                  3/24,2/24,1/24))


kable(PMF, align = "lc")
```

|      |      |      |      |      |      |      |      |      |      |
|------|------|------|------|------|------|------|------|------|------|
| x    | 2    | 3    | 4    | 5    | 6    | 7    | 8    | 9    | 10   |
| p(x) | 1/24 | 2/24 | 3/24 | 4/24 | 4/24 | 4/24 | 3/24 | 2/24 | 1/24 |

Use your probability mass function to determine the following probabilities:

-   P(X = 4) = **3/24**

```{r}
cat("P(X = 4) = 3/24 ", "or", PMF$Probability[PMF$Results == 4], "\n")
```

-   P(X = 12) = 0/24 =**0**
-   P(X is Even) = **12/24**
-   P(X = 4 or X = 5) = $P(4 \cup 5)$ = 3/24 + 4/24 = **7/24**

#### B

Determine the expected value and variance of the probability distribution (PMF) of the sum of one six-sided die and one four-sided die \[1-4\].

Expected Value:

```{r}

E = (1/24)*2 + (2/24)*3 + (3/24)*4 + (4/24)*5 + (4/24)*6 + 
  (4/24)*7 + (3/24)*8 + (2/24)*9 + (1/24)*10
E
```

Variance:

```{r}

V = (1/24)*(2-E)^2 + (2/24)*(3-E)^2 + (3/24)*(4-E)^2 + (4/24)*(5-E)^2 + 
  (4/24)*(6-E)^2 + (4/24)*(7-E)^2 + (3/24)*(8-E)^2 + 
  (2/24)*(9-E)^2 + (1/24)*(10-E)^2

V
 

```

**Using vectors:**

```{r}

Expected <- sum(PMF$Results*PMF$Probability)
cat("E[X] =",Expected, "\n")

Variance <- sum(((PMF$Results - Expected)^2)*PMF$Probability)
cat("V[X] =",Variance, "\n")


```

\newpage

#### C

What is the cumulative distribution function (CDF) of the sum of one six-sided die and one four-sided die \[1-4\]?

|      |      |      |      |       |       |       |       |       |       |
|------|------|------|------|-------|-------|-------|-------|-------|-------|
| x    | 2    | 3    | 4    | 5     | 6     | 7     | 8     | 9     | 10    |
| p(x) | 1/24 | 2/24 | 3/24 | 4/24  | 4/24  | 4/24  | 3/24  | 2/24  | 1/24  |
| F(X) | 1/24 | 3/24 | 6/24 | 10/24 | 14/24 | 18/24 | 21/24 | 23/24 | 24/24 |

```{r}

#OR

PMF$CDF <- cumsum(PMF$Probability)
kable(PMF, align = "lcc")

```

```{r}
Probability = c(1/24,2/24,3/24,4/24,4/24,4/24,
                                  3/24,2/24,1/24)

cumsum(Probability)
```

Use the CDF to determine the following probabilities:

-   P(X $\leq$ 6) = **14/24**

```{r}
# Or

cat("P(X <= 6)=",PMF$CDF[PMF$Results == 6], "\n")
```

-   P(X \> 3) = 1 - P(X $\leq$ 3) = **21/24**
-   P(5 $\leq$ X $\leq$ 7) = P(X $\leq$ 7) - P(X $\leq$ 4) = 18/24 - 6/24 = **12/24**
-   P(X $\leq$ 3 or X \> 7) = P(X $\leq$ 3) + P(X \> 7) = 3/24 + 6/24 = **9/24**

## Task 2

#### A

Suppose we are rolling a single six-sided die. We define a success to be landing on a 5.

-   Define the Bernoulli distribution for this experiment.
-   Calculate the expected value and variance (from theory).

```{=latex}
\begin{itemize}
      \item The PMF of the Bernoulli distribution is:
      \end{itemize}
      \begin{align}
p(x) =\left\{
\begin{array}{ll}
      p, & \text{if~} x = \text{5} \\
      1-p, & ,~\text{Otherwise} \\
\end{array} 
\right.
    \end{align} 
    where \textbf{$p$ = 1/6 is the probability of landing on a 5. Therefore:}
    \begin{itemize}
        \item Expected value: $p$ = \textbf{1/6}
        \item Variance: $p(1-p)$ = \textbf{5/36 = 0.1388889}
    \end{itemize}
```

#### B

Suppose we repeat the experiment from A *n=12* times.

-   What is the name of the resulting distribution?

    -   **Binomial distribution**

-   What is the expected value and variance of this distribution?

    -   $E[X]$ = $n\times p$ = 12\$\times\$1/6 = **2**
    -   $V[X]$ = 12 $\times$ $p(1-p)$ = 12 $\times$ 5/36 = **5/3 = 1.666667**

-   Calculate the probability that we get 5 successes.

    ```{r}

    cat("P(X = 5) =",dbinom(5, size = 12, prob = 1/6) ,"\n")
    ```

-   Calculate the probability that we get more than 5 successes.

    ```{r}

    cat("P(X > 5) =",1 - pbinom(5, size = 12, prob = 1/6, 
                                lower.tail = TRUE) ,"\n")

    ```

#### C

Suppose we have a Poisson distribution for the number of arrivals to our restaurant with $\lambda$= 7 per hour.

-   What is the expected value and variance of this distribution?

    $E[X] = \lambda = 7$\
    $V[X] = \lambda = 7$

-   What is the probability that we have *exactly* 5 customers in an hour?

```{r}

cat("P(X = 5) =",dpois(5,lambda = 7) ,"\n")

```

-   What is the probability that we have *at most* 8 customers in an hour?

```{r}

cat("P(X <= 8) =",ppois(8,lambda = 7) ,"\n")


```

-   What is the probability that we have *more than* 8 customers in an hour?

```{r}

cat("P(X > 8) =",1 - ppois(8,lambda = 7) ,"\n")

# OR

cat("P(X > 8) =",ppois(8,lambda = 7,lower.tail = FALSE) ,"\n")

```

\newpage

## Task 3

#### A

Assume that $X$ follows a Uniform(a = 1, b = 3) distribution and $Y$ follows a Uniform(a = 0.5, b = 10.5). Calculate the following probabilities:

-   P(X \< 3.5)

    ```{r}

    cat("P(X < 3.5) =",punif(3.5,min = 1,max = 3,lower.tail = TRUE) ,"\n")

    ```

-   P(Y \< 3.5)

    ```{r}

    cat("P(Y < 3.5) =",punif(3.5,min = 0.5,max = 10.5,lower.tail = TRUE) 
        ,"\n")

    ```

-   P(X \> 2.5)

    ```{r}

    cat("P(X > 2.5) =",punif(2.5,min = 1,max = 3,lower.tail = FALSE) 
        ,"\n")

    ```

-   P(Y \> 2.5)

    ```{r}

    cat("P(Y > 2.5) =",punif(2.5,min = 0.5,max = 10.5,
                             lower.tail = FALSE) ,"\n")

    ```

-   P(2.3 \< X \< 3)

    ```{r}

    cat("P(2.3 < X < 3) =",punif(2.3,min = 1,max = 3,
                                 lower.tail = FALSE) ,"\n")

    ```

-   P(2.3 \< Y \< 5.3)

    ```{r}

    cat("P(2.3 < Y < 5.3) =",punif(5.3,min = 0.5,max = 10.5,
                                   lower.tail = TRUE) -
    punif(2.3,min = 0.5,max = 10.5,lower.tail = TRUE) ,"\n")

    ```

#### B

Assume that $Z$ follows a standard Normal($\mu$ = 0, $\sigma$ = 1) distribution and $T$ follows a Normal($\mu$ = 3, $\sigma$ = 2). Calculate the following probabilities:

-   P(Z \< 1)

    ```{r}

    cat("P(Z < 1) =", pnorm(1,mean=0,sd=1,lower.tail = TRUE)
        ,"\n")

    ```

-   P(T \< 1)

    ```{r}

    cat("P(T < 1) =", pnorm(1,mean=3,sd=2,lower.tail = TRUE)
        ,"\n")

    pnorm(1,mean=3,sd=2)

    ```

-   P(-0.5 \< Z \< 0.5)

    ```{r}

    cat("P(-0.5 < Z < 0.5) =", pnorm(0.5,mean=0,sd=1,
                                     lower.tail = TRUE) -
    pnorm(-0.5,mean=0,sd=1,lower.tail = TRUE),"\n")

    ```

-   P(-0.5 \< T \< 0.5)

    ```{r}

    cat("P(-0.5 < T < 0.5) =", pnorm(0.5,mean=3,sd=2,
                                     lower.tail = TRUE) -
    pnorm(-0.5,mean=3,sd=2,lower.tail = TRUE),"\n")

    pnorm(0.5,mean=3,sd=2) - pnorm(-0.5,mean=3,sd=2)

    ```
