---
title: "STAT1200 Continuous Probability Distributions Examples"
author: "Sean Hellingman"
format: pdf
editor: visual
---

## Example 1 from slides

#### Single Integration

```{r}


my_distribution <- function(x){
  
  out <- 2 - 2*x
  
  }

integrate(my_distribution,
          lower = 0,
          upper = 1)



integrate(my_distribution,
          lower = 0.5,
          upper = 0.5)


```

```{r definition}

curve(expr = my_distribution, from = -0.2, to = 1.2 )
abline(v = 0)
abline(h = 0)
```

## Example 2 from slides

```{r Graphics}

par(mfrow = c(1, 2))

curve(expr = my_distribution, from = -0.2, to = 1.2 )
abline(v = c(0,0.2),col=c("black","red"))
abline(h = 0)



curve(expr = my_distribution, from = -0.2, to = 1.2 )
abline(v = c(0,0.1,0.3),col=c("black","red","red"))
abline(h = 0)

```

We could use the areas of right triangles to arrive at the solutions: Area = $(a\cdot b)/2$

1.  $P(X<0.2) = 1 - P(X >= 0.2)$

$f(x) = 2 -2x$

$f(0.2) = 2 - 2*(0.2) = 1.6 = a$

$1 - 0.2 = 0.8 = b$

$1- (0.8)(1.6)/2 = 1 - 0.64 = 0.36$

2.  $P(0.1 < X < 0.3)$

$f(0.1) = 2 - 2(0.1) = 1.8 = a1$

$f(0.3) = 2 - 2(0.3) = 1.4 = a2$

$b1 = 1 - 0.1 = 0.9 b2 = 1 - 0.3 = 0.7$

$A1 = (1.8)(0.9)/2 = 0.81$

$A2 = (1.4)(0.7)/2 = 0.49$

$P(0.1 < X < 0.3) = 0.81 - 0.49 = 0.32$

Or, we could use calculus (works in general)

```{r Calculus}

# P(X < 0.2)

integrate(my_distribution,
         lower = 0,
         upper = 0.2)

# P(0.1 < X 0.3)

integrate(my_distribution,
         lower = 0.1,
         upper = 0.3)


```

Plot standard normal distribution:

```{r Normal plot}


# Create a sequence of numbers between -10 and 10 incrementing by 0.01.
x <- seq(-10, 10, by = .01)

# Standard normal
y <- dnorm(x, mean = 0, sd = 1)


plot(x,y,main = "Standard Normal Distribution",xlim = c(-4,4))



```

## Example 3 from slides

```{r 4-plots}

x <- seq(-30, 30, by = .01)

par(mfrow = c(1, 2))

## Uniform

unif1 <- dunif(x,min=-2,max=1)
plot(x,unif1,main = "Uniform(-2,1)", xlim = c(-5,5),cex=0.25)
abline(v = c(-2,1))
abline(h = 0)

unif2 <- dunif(x,min=1,max=10)
plot(x,unif2,main = "Uniform(1,10)", xlim = c(0,15),cex=0.25)
abline(v = c(1,10))
abline(h = 0)

## Normal

Norm2 <- dnorm(x, mean = 1, sd = 1)
Norm3 <- dnorm(x, mean = -2, sd = 6)

plot(x,Norm2,main = "Normal Dist. mean:1 sd:1",xlim = c(-30,30))
plot(x,Norm3,main = "Normal Dist. mean:-2 sd:6",xlim = c(-30,30))
```

CDFs

```{r}

# Create a sequence of numbers between -20 and 20 incrementing by 0.001.
x <- seq(-20, 20, by = .001)

par(mfrow = c(2, 2))

# Uniform distributions

unif1pdf <- dunif(x,min=0,max=1)
unif1cdf <- punif(x,min =0, max=1)

plot(x,unif1pdf,main = "PDF of Uniform(0,1)", xlim = c(-0.1,1.1),ylab = "f(x)",cex=0.25)
abline(v = c(0,1))
abline(h = 0)
 plot(x,unif1cdf,main = "CDF of Uniform(0,1)", xlim = c(-0.1,1.1),ylab = "F(x)")

## Normal ##


Norm1pdf<- dnorm(x, mean = 0, sd = 1)
Norm1cdf <- pnorm(x, mean = 0, sd = 1)

plot(x,Norm1pdf,main = "PDF of Standard Normal Distribution",xlim = c(-4,4),ylab = "f(x)", cex.main=0.75)
plot(x,Norm1cdf,main = "CDF of Standard Normal Distribution",xlim = c(-4,4),ylab = "F(x)", cex.main=0.75)


```

## Example 4 from slides

```{r}

#pnorm(0.5,mean=0,sd=1,lower.tail = TRUE)
#pnorm(1,mean = 0, sd = 1, lower.tail = FALSE)

# Create a sequence of numbers between -10 and 10 incrementing by 0.01.
x <- seq(-10, 10, by = .01)

# Standard normal
y <- dnorm(x, mean = 0, sd = 1)


plot(x,y,main = "Standard Normal Distribution",xlim = c(-4,4))
abline(v = 0.5,col="red")
abline(h = 0)

#P(X<0.5):
pnorm(0.5,mean=0,sd=1,lower.tail = TRUE)

plot(x,y,main = "Standard Normal Distribution",xlim = c(-4,4))
abline(v = c(-0.1,0.1),col=c("red","red"))
abline(h = 0)

#P(-0.1 < X < 0.1):
pnorm(0.1,lower.tail = TRUE) - pnorm(-0.1,lower.tail = TRUE)

plot(x,y,main = "Standard Normal Distribution",xlim = c(-4,4))
abline(v = 1,col="blue")
abline(h = 0)

#P(X>1)

pnorm(1,mean = 0, sd= 1, lower.tail = FALSE)

# Or 

1 - pnorm(1,mean = 0, sd= 1, lower.tail = TRUE)
```

Plot exponential distribution:

```{r exp plot}


# Create a sequence of numbers between -10 and 10 incrementing by 0.01.
x <- seq(0, 10, by = .01)

# Standard normal
y <- dexp(x, rate = 1)


plot(x,y,main = "Exponential Distribution",xlim = c(-0.05,4))


## Probabilites ##


# P(X<2)

pexp(2,rate = 1,lower.tail = TRUE)

# P(X>4)

pexp(4,rate = 1,lower.tail = FALSE)
```

## Example 5 from slides

Food and Housing prices go up:

$P(X=1,Y=1)$ = 0.25

Food prices are unchanged (ignoring Housing prices):

$P(X=0)$ = 0.18
