---
title: "STAT1200 Linear Regression II Examples"
author: "Sean Hellingman"
format: pdf
editor: visual
---

### Data Description:

A data.frame with 935 observations on 17 variables:

-   wage: monthly earnings
-   hours: average weekly hours
-   IQ: IQ score
-   KWW: knowledge of world work score
-   educ: years of education
-   exper: years of work experience
-   tenure: years with current employer
-   age: age in years
-   married: =1 if married
-   black: =1 if black
-   south: =1 if live in south
-   urban: =1 if live in SMSA (standard metropolitan statistical area)
-   sibs: number of siblings
-   brthord: birth order
-   meduc: mother's education (years)
-   feduc: father's education (years)
-   lwage: natural log of wage

```{r}

library(MASS)
library(dplyr)
data(Cars93)
colnames(Cars93)

# Visualize

Cars93 %>% 
  dplyr::select(Price, MPG.highway,
                EngineSize, Horsepower,
                Weight) %>% 
                pairs()

# Individual Models

lmA <- lm(Price ~ MPG.highway, data = Cars93)
summary(lmA)


lmB <- lm(Price ~ EngineSize, data = Cars93)
summary(lmB)


lmC <- lm(Price ~ Horsepower, data = Cars93)
summary(lmC)


lmD <- lm(Price ~ Weight, data = Cars93)
summary(lmD)

# Multiple Linear Regression

lmABCD <- lm(Price ~ MPG.highway + EngineSize +
             Horsepower + Weight, data = Cars93)
summary(lmABCD) 

```

# Example 1 from slides

```{r}


wage2 <- read.csv("wage2.csv")


```

# Example 2 from slides

```{r}

summary(lmABCD)
# Almost 62% of the variability


```

# Example 3 from slides

```{r}

summary(lmABCD)
# ANOVA: at least one is significantly different than zero
# Only Horsepower is non-zero


```

# Example 4 from slides

```{r}

library(car)
vif(lmABCD) # problem with weight
barplot(vif(lmABCD), main = "VIF Values", horiz = TRUE,
col = "steelblue",bty = "l",
las = 1,cex.names = .55)
abline(v = 5, lwd = 3, lty = 2)



```

# Example 5 from slides

```{r}



# Linearity 

plot(lmABCD$residuals)
abline(h=0,col="blue")

plot(lmABCD,1)


# Normality 

plot(lmABCD,2) # Q-Q plot
lmABCD.standard <- rstandard(lmABCD) #standardized residuals

# Reject the null hypothesis of normality
shapiro.test(lmABCD.standard)



# Constant Variance (Homoscedasticity)

plot(lmABCD,3) #Not okay

ncvTest(lmABCD) #pass the test (Reject the null)



# Independence

lmABCD.standard <- rstandard(lmABCD) #standardized residuals

plot(lmABCD.standard)
abline(h=0,col="blue") # Almost Looks okay


durbinWatsonTest(lmABCD) # Reject null hypothesis of independence 

vif(lmABCD) # Variance Inflation Factors
# REMOVE Weight

lmABC <- lm(Price ~ MPG.highway + EngineSize +
             Horsepower, data = Cars93)
summary(lmABC) 

### IN PRACTICE WE WILL NEED TO RE-ASSES THE MODELS ###
```
