---
title: "STAT1200 Sampling and Estimation Examples"
author: "Sean Hellingman"
format: pdf
editor: visual
---

#### Packages

```{r packages}
rm(list = ls()) #Remove all objects from the workspace

if(!require(BSDA)) install.packages("BSDA")
library(BSDA)
if(!require(DescTools)) install.packages("DescTools")
library(DescTools)

```

Assume a population standard deviation of 68 for *hp*.

```{r cars-CI}

#Using formula from slides:

mean.hp <- mean(mtcars$hp)

Lower.95 <- mean.hp - qnorm(p=(0.05/2),lower.tail = FALSE)*(68/sqrt(nrow(mtcars)))
Upper.95 <- mean.hp + qnorm(p=(0.05/2),lower.tail = FALSE)*(68/sqrt(nrow(mtcars)))

#Using the BSDA package:
?z.test()
HP.CI <- z.test(mtcars$hp,sigma.x = 68,conf.level = 0.95)
HP.CI$conf.int



```

## Example 1 from slides

```{r CI-ex1}

data("mtcars")

xbar = mean(mtcars$mpg)

n = nrow(mtcars)

# Lower.95 <- xbar - 0.69847*(6/sqrt(n))
# Upper.95 <- mean.hp + 0.69847*(6/sqrt(n))


# 95% C.I.

Ex1.1 <- z.test(mtcars$mpg,sigma.x = 6,conf.level = 0.95)
Ex1.1$conf.int


# 99% C.I.

Ex1.2 <- z.test(mtcars$mpg,sigma.x = 6,conf.level = 0.99)
Ex1.2$conf.int
```

Finite population correction factor:

```{r}
set.seed(123)
library(dplyr)


random.flowers <- iris %>%
  sample_n(10, replace = FALSE)
random.flowers

petal.width.mean <- mean(random.flowers$Petal.Width)

Lower.95 <- petal.width.mean - qnorm(p=0.05/2,lower.tail=FALSE)*(sd(iris$Petal.Width)/sqrt(length(random.flowers)))*
  sqrt((nrow(iris) - nrow(random.flowers))/(nrow(iris)-1))

Upper.95 <- petal.width.mean + qnorm(p=0.05/2,lower.tail=FALSE)*(sd(iris$Petal.Width)/sqrt(length(random.flowers)))*
  sqrt((nrow(iris) - nrow(random.flowers))/(nrow(iris)-1))
  
  
Results <- list(Lower=Lower.95,Mean=petal.width.mean,Upper=Upper.95)
Results
```

## Example 2 from slides

```{r FPC}




```

Plotting distributions

```{r}


# Load the ggplot2 library
library(ggplot2)

# Set the degrees of freedom for the t-distributions
df1 <- 3  # Degrees of freedom for the first t-distribution
df2 <- 10 # Degrees of freedom for the second t-distribution

# Generate data for the standard normal distribution
x <- seq(-4, 4, length=1000)
std_normal <- dnorm(x)

# Generate data for the two t-distributions
t_dist1 <- dt(x, df = df1)
t_dist2 <- dt(x, df = df2)

# Create a data frame for plotting
data <- data.frame(x = x, std_normal = std_normal, t_dist1 = t_dist1, t_dist2 = t_dist2)

# Create the plot
plot <- ggplot(data, aes(x)) +
  geom_line(aes(y = std_normal, color = "Standard Normal"), size = 1) +
  geom_line(aes(y = t_dist1, color = "t-distribution (df=3)"), size = 1) +
  geom_line(aes(y = t_dist2, color = "t-distribution (df=10)"), size = 1) +
  labs(
    title = "Standard Normal Distribution vs. t-distributions",
    x = "x",
    y = "Density"
  ) +
  theme_minimal() +
  scale_color_manual(
    values = c("Standard Normal" = "blue", "t-distribution (df=3)" = "red", "t-distribution (df=10)" = "green")
  )

# Display the plot
print(plot)


```

### Unknown sigma

```{r hp}

#Using formula from slides:

?qt()

mean.hp <- mean(mtcars$hp)

Lower.95.A <- mean.hp - qt(0.05/2,df=nrow(mtcars)-1,lower.tail = FALSE)*(sd(mtcars$hp)/sqrt(nrow(mtcars)))

Upper.95.A <- mean.hp + qt(0.05/2,df=nrow(mtcars)-1,lower.tail = FALSE)*(sd(mtcars$hp)/sqrt(nrow(mtcars)))


#Using the t.test
?t.test()

HP.CI.A <- t.test(mtcars$hp,conf.level = 0.95)
HP.CI.A$conf.int




```

## Example 3 from slides

```{r petals-unknown-sigma}




```

### proprtions

Proportion of cars with 6 cylinders

```{r cars-cyl}

#Calculate the sample proportion
cyl.prop <- sum(as.numeric(mtcars$cyl == 6))/nrow(mtcars)

cyl.prop.lower.90 <- cyl.prop - qnorm(p=0.10/2,lower.tail=FALSE)*sqrt((cyl.prop*(1-cyl.prop))/nrow(mtcars))

cyl.prop.upper.90 <- cyl.prop + qnorm(p=0.10/2,lower.tail=FALSE)*sqrt((cyl.prop*(1-cyl.prop))/nrow(mtcars))


#Or using prop.test()
?prop.test()

cyl.prop.test <- prop.test(sum(as.numeric(mtcars$cyl == 6)),nrow(mtcars), conf.level=0.9) 
cyl.prop.test$conf.int

```

## Example 4 from slides

```{r fav-colours}

Colours <- c('red', 'red', 'blue', 'green', 'green', 'orange', 'red', 'orange', 'yellow', 'green', 'red', 'orange')





```

### Variance

```{r chi-squared}


# defining degrees of freedom
df = 5
  
# calculating for the values in the interval [0,5]
print ("Calculating for the values [0,5]")
pchisq(5, df = df,lower.tail = TRUE)
  
# calculating for the values in the interval [5,inf)
print ("Calculating for the values [5,inf)")
pchisq(5, df = df,lower.tail = FALSE)


```

```{r chi-squared-plot}



# Create a sequence of x values
x <- seq(0, 20, length.out = 1000)

# Create a plot with 8 chi-squared densities
plot(x, dchisq(x, df = 1), type = "l", col = "red", lwd = 2, ylim = c(0, 0.5), 
     xlab = "x", ylab = "Density", main = "Chi-Squared Distributions")

for (df in 2:8) {
  lines(x, dchisq(x, df = df), col = rainbow(8)[df], lwd = 2)
}

# Add a legend
legend("topright", legend = c("df = 1", "df = 2", "df = 3", "df = 4", "df = 5", "df = 6", "df = 7", "df = 8"),
       col = rainbow(8), lwd = 2, cex = 0.8)


```

```{r var-CI}


var(mtcars$hp)

VarCI(mtcars$hp,method = "classic",conf.level = 0.95)



```

## Example 5 from slides

```{r var-CI-example}


# VarCI(iris$Petal.Length,method = "classic", conf.level = 0.95) # Use this one!!



```

Sample sizes

```{r sample-sizes}

library(EnvStats)

sigma <- 1.2
E <- 0.5


#Own code
n_needed <- ((qnorm(p=0.05/2,lower.tail=FALSE)*sigma)/E)^2
n_needed

#Using function
ciNormN(E,sigma,conf.level = 0.95) # Use this one!



```

## Example 6 from slides

```{r}


E <- 0.25
sigma <- 3.3
alpha <- 0.05 #=> C.I. is 95%

n_req <- ((qnorm(p=alpha/2,lower.tail=FALSE)*sigma)/E)^2

n_req

ciNormN(E,sigma,conf.level = 0.95)



```

Differences in average hp between cars with 4 and 6 cylinders

```{r hp-cyl}

library(DescTools)

data("mtcars")

MeanDiffCI(mtcars$hp[mtcars$cyl == 4],mtcars$hp[mtcars$cyl == 6],conf.level = 0.95,
  paired = FALSE)



```

## Example 7 from slides

```{r petal-lenght-diff}

data("iris")

MeanDiffCI(iris$Petal.Length[iris$Species == "setosa"],iris$Petal.Length[iris$Species == "versicolor"],conf.level = 0.90, paired = FALSE, method = "classic")


```

## Example 8 from slides

```{r}

?prop.test()

x1 <- 0.25*52
x2 <- 0.35*60

n1 <- 52
n2 <- 60


CI.Prop.diff<- prop.test(x=c(x1,x2),n=c(n1,n2),conf.level = 0.95,correct=FALSE)


CI.Prop.diff$conf.int




```
